---
title: "CIND 820 Big Data Final Project"
author: "Jossa Soto"
date: "06/03/2022"
output:
  html_document:
    keep_md: yes
  word_document: default
  pdf_document: default
---

```{r setup, message=FALSE}
# Load needed libraries
library(tidyverse)
library(ROSE)
library(tidytext)
library(glue)
library(stringr)
library(ggplot2)
library(patchwork)
library(gt)
library(tm)
library(knitr)
library(NLP)
library(wordcloud)
library(RColorBrewer)
library(rpart)
library(rpart.plot)
library(randomForest)
library(caTools)
library(caret)
library(e1071)
library(MASS)
```

## Dataset
```{r}
twitter_balanced <- read.csv("Twitter US Airline Sentiment.csv", header = T, na.strings = c("","NA"))

table(twitter_balanced$airline_sentiment)
```
Data is imbalanced so sampling will be done so that the subset is balanced. This will be done by undersampling the negative and neutral group. 

```{r}
twitter_balanced$airline_sentiment <- as.factor(twitter_balanced$airline_sentiment)

twitter_pos_ind <- which(twitter_balanced$airline_sentiment == "positive")
twitter_neg_ind <- which(twitter_balanced$airline_sentiment == "negative")
twitter_neut_ind <- which(twitter_balanced$airline_sentiment == "neutral") 

set.seed(2407)
nsample <- 2500
pick_neg <- sample(twitter_neg_ind, nsample)
pick_neut <- sample(twitter_neut_ind, nsample)

twitter_balanced <- twitter_balanced[c(pick_neg, pick_neut, twitter_pos_ind),] 

table(twitter_balanced$airline_sentiment)
```

```{r}
# Declare each variable from df for easy use
tweet_id <- twitter_balanced$tweet_id
airline_sentiment <- twitter_balanced$airline_sentiment
airline_sentiment_confidence <- twitter_balanced$airline_sentiment_confidence
negativereason <- twitter_balanced$negativereason
negativereason_confidence <- twitter_balanced$negativereason_confidence
airline <- twitter_balanced$airline
airline_sentiment_gold <- twitter_balanced$airline_sentiment_gold
name <- twitter_balanced$name
negativereason_gold <- twitter_balanced$negativereason_gold
retweet_count <- twitter_balanced$retweet_count
tweet_text <- twitter_balanced$text
tweet_coord <- twitter_balanced$tweet_coord
tweet_created <- twitter_balanced$tweet_created
tweet_location <- twitter_balanced$tweet_location
user_TZ <- twitter_balanced$user_timezone

# Display first 6 rows of data
head(twitter_balanced)
```

## Data Exploration

```{r}
str(twitter_balanced)

# Change airline_sentiment to factors
twitter_balanced$airline <- as.factor(twitter_balanced$airline)
twitter_balanced$airline_sentiment <- as.factor(twitter_balanced$airline_sentiment)
twitter_balanced$airline_sentiment_gold <- as.factor(twitter_balanced$airline_sentiment_gold)
twitter_balanced$negativereason <- as.factor(twitter_balanced$negativereason)
twitter_balanced$negativereason_gold <- as.factor(twitter_balanced$negativereason_gold)

# Change tweet_created from char to date
twitter_balanced$tweet_created <- as.Date(twitter_balanced$tweet_created)

# Rerun str and summary to display updated variables
str(twitter_balanced)

# Display summary of the twitter_balanced data frame
summary(twitter_balanced)

# Standard deviation for numerical variables
sd(airline_sentiment_confidence); sd(!is.na(negativereason_confidence)); sd(retweet_count)
```

```{r}
# Check for and count any NAs
# Unlist to convert the list output to a vector for easier readability
unlist(lapply(lapply(twitter_balanced, is.na), sum))
```
```{r warning = FALSE}
# Check for any outliers in the numerical variables

# Airline sentiment confidence
ASC_box <- twitter_balanced %>% ggplot() + 
  geom_boxplot(aes(y = airline_sentiment_confidence), outlier.colour="black", outlier.shape=16, outlier.size=2, notch=FALSE) 

# Negative reason confidence
NRC_box <- twitter_balanced %>% 
  ggplot() + 
  geom_boxplot(aes(y = negativereason_confidence), outlier.colour="black", outlier.shape=16, outlier.size=2, notch=FALSE) 

# Retweet count
RC_box <- twitter_balanced %>% 
  ggplot() + 
  geom_boxplot(aes(y = retweet_count), outlier.colour="black", outlier.shape=16, outlier.size=2, notch=FALSE)

outlier_fig <- ASC_box + NRC_box + RC_box + plot_layout(ncol = 3)
outlier_fig
```

```{r}
# Frequency of the factor and discrete variables
table(airline)
table(airline_sentiment)
table(negativereason)
table(negativereason_gold)
table(airline_sentiment_gold)
table(retweet_count)
```

```{r}
# Airline
airline_bar <- twitter_balanced %>% ggplot() + 
  geom_bar(aes(x = airline, fill = airline)) +
  theme(axis.text.x = element_blank()) +
  ggtitle("Airline") + labs(y = "Count", x = "Airline", fill = "Airline")

# Airline sentiment
AS_bar <- twitter_balanced %>% ggplot() + 
  geom_bar(aes(x = airline_sentiment, fill = airline_sentiment)) +
  theme(axis.text.x = element_blank()) +
  ggtitle("Airline Sentiment (Class)") + labs(y = "Count", x = "Airline Sentiment (Class)", fill = "Airline Sentiment (Class)")

# Airline sentiment
ASG_bar <- twitter_balanced %>% ggplot() + 
  geom_bar(aes(x = airline_sentiment_gold, fill = airline_sentiment_gold)) +
  theme(axis.text.x = element_blank()) +
  ggtitle("Airline Sentiment - Gold") + labs(y = "Count", x = "Airline Sentiment - Gold", fill = "Airline Sentiment - Gold")

# Negative reason
NR_bar <- twitter_balanced %>% ggplot() + 
  geom_bar(aes(x = negativereason, fill = negativereason)) +
  theme(axis.text.x = element_blank()) +
  ggtitle("Negative Reason") + labs(y = "Count", x = "Negative Reason", fill = "Negative Reason")

# Negative reason gold
NRG_bar <- twitter_balanced %>% ggplot() + 
  geom_bar(aes(x = negativereason_gold, fill = negativereason_gold)) +
  theme(axis.text.x = element_blank()) +
  ggtitle("Negative Reason - Gold") + labs(y = "Count", x = "Negative Reason - Gold", fill = "Negative Reason - Gold")

frequency_fig1 <- airline_bar + AS_bar + ASG_bar + NRG_bar + plot_layout(ncol = 2, nrow = 2)
frequency_fig1
NR_bar
```

## Preprocessing

Before preproccessing, a corpus needs to be created from the "text" variable of the dataset. 

```{r, warning = FALSE}
tweets_text <- twitter_balanced$text
#tweets_text <- boost_tokenizer(twitter_balanced$text)
str(tweets_text)

# Create a corpus
tweets_source <- VectorSource(tweets_text)

tweets_corpus <- Corpus(tweets_source)

tweets_corpus <- tm_map(tweets_corpus, function(x) iconv(enc2utf8(x), sub = "byte"))
```
Preprocessing is a multi-step process where extraneous elements including punctuation, number, and stop words are removed, and clean the data to minimize the words that could create noise to the data and reduce the model down to the most important words.  

```{r, warning = FALSE}
tweets_corpus <- tm_map(tweets_corpus, content_transformer(tolower)) # Convert all text to lowercase

tweets_corpus <- tm_map(tweets_corpus, removePunctuation) # Remove all punctuation

tweets_corpus <- tm_map(tweets_corpus, removeNumbers) # Remove numbers

# Custom function to remove extra elements/characters from the text
textprocessing <- function(x){
  gsub("http[[:alnum:]]*",'', x)
  gsub('http\\S+\\s*', '', x) # Remove URLs
  gsub('\\b+RT', '', x) # Remove RT
  gsub('#\\S+', '', x) # Remove Hashtags
  gsub('@\\S+', '', x) # Remove Mentions
  gsub('[[:cntrl:]]', '', x) # Remove Controls and special characters
  gsub("\\d", '', x) # Remove Controls and special characters
  gsub('[[:punct:]]', '', x) # Remove Punctuations
  gsub("^[[:space:]]*","",x) # Remove leading whitespaces
  gsub("[[:space:]]*$","",x) # Remove trailing whitespaces
  gsub(' +',' ',x) # Remove extra whitespaces 
}

tweets_corpus <- tm_map(tweets_corpus, textprocessing)

tweets_corpus <- tm_map(tweets_corpus, stripWhitespace) # Remove whitespace

# Create custom list of stop words
stop_words <- c(stopwords("english"), "rt","southwestair", "americanair", "delta", "united", "usairway", "virginamerica", "jetblue", "amp")

tweets_corpus <- tm_map(tweets_corpus, removeWords, stop_words) # Remove stop words

# Create a copy of tweets pre-stemming
tweets_corpus_copy <- tweets_corpus

tweets_corpus <- tm_map(tweets_corpus, stemDocument) # Stem the text

# Display a row of the cleaned corpus
tweets_corpus[[148]][1]
```
## Feature Selection 

#### Document-Term: TF-IDF Weights

Create a document-term matrix with TF-IDF weights then reduce the matrix down so it only contains a certain \% of the terms. 

Weights are calculated using the following formula:    

$TF-IDF(t)=TF(t) \times IDF(t)$
\n $w_{i,j} = tf_{i,j} \times log\frac{N}{df_i}$

```{r, warning = FALSE}
tweets_dtm_tfidf <- DocumentTermMatrix(tweets_corpus, control = list(weighting = weightTfIdf))
tweets_dtm_tfidf

tweets_m_tfidf <- as.matrix(tweets_dtm_tfidf)
dim(tweets_m_tfidf)

# Display a subset of the matrix
tweets_m_tfidf[14:28, 10:15]

# Reduce the matrix by removing the low frequency terms
tweets_dtm_tfidf_sparse <- removeSparseTerms(tweets_dtm_tfidf, 0.99) 
tweets_dtm_tfidf_sparse

tweets_m_tfidf <- as.matrix(tweets_dtm_tfidf_sparse)
dim(tweets_m_tfidf)

# Display a subset of the matrix after the low frequency terms were removed
tweets_m_tfidf[14:28, 10:15]
```


#### Bag of Words

A document-term matrix is created with Bag-of-Words applied. For this model, the corpus is only checked whether the words exist, not where nor the context. 

The DTM is also reduced to remove extraneous words. 

```{r}
tweets_dtm_bow <- DocumentTermMatrix(tweets_corpus)
tweets_dtm_bow

tweets_m_bow <- as.matrix(tweets_dtm_bow)
dim(tweets_m_bow)

# Display a subset of the matrix
tweets_m_bow[14:28, 10:15]

# Reduce the matrix by removing the low frequency terms
tweets_dtm_bow_sparse <- removeSparseTerms(tweets_dtm_bow, 0.99) 
tweets_dtm_bow_sparse

tweets_m_bow <- as.matrix(tweets_dtm_bow_sparse)
dim(tweets_m_bow)

# Display a subset of the matrix after the low frequency terms were removed
tweets_m_bow[14:28, 10:15]
```
#### Unigrams (try bigram + trigram as well)

Unigrams is a one-term version of n-gram model which predicts the occurrence of the words based on the occurrence of its n-1 previous words. 

The DTM is also reduced to remove extraneous words. 

```{r}
Unigramtokenizer <- function(x) {
        unlist(lapply(ngrams(words(x), 1), paste, collapse = " "), use.names = FALSE)}
```

```{r, warning = FALSE}
# Unigram dtm
tweets_dtm_uni <- DocumentTermMatrix(tweets_corpus,control = list(tokenize = Unigramtokenizer))

tweets_m_uni <- as.matrix(tweets_dtm_uni)
dim(tweets_m_uni)

# Display a subset of the matrix
tweets_m_uni[14:28, 10:15]

# Reduce the matrix by removing the low frequency terms
tweets_dtm_uni_sparse <- removeSparseTerms(tweets_dtm_uni, 0.99) 
tweets_dtm_uni_sparse

tweets_m_uni <- as.matrix(tweets_dtm_uni_sparse)
dim(tweets_m_uni)

# Display a subset of the matrix after the low frequency terms were removed
tweets_m_uni[14:28, 10:15]
```

## Classifier

### Decision Tree

Apply the Decision Tree algorithm to the three models: TF-IDF, Bag-of-Words, and Unigram. 

#### TF-IDF

```{r}
m_tfidf <- as.data.frame(tweets_m_tfidf)
colnames(m_tfidf) <- make.names(colnames(m_tfidf))
m_tfidf$class = twitter_balanced$airline_sentiment
``` 

First attempt (no repetition) 
```{r}
set.seed(439) #use repeated k-fold validation to repeat then record 20 accuracy scores in a table to proper compare the three classifiers (parametric/non-parametric tests on the mean)

split_tfidf = sample.split(m_tfidf$class, SplitRatio = 0.7)
train_tfidf = subset(m_tfidf, split_tfidf == TRUE)
test_tfidf = subset(m_tfidf, split_tfidf == FALSE)

DT_model_tfidf <- rpart(class~., data = train_tfidf, method = "class")

predictDT_tfidf <- predict(DT_model_tfidf, newdata = test_tfidf, type = "class")

summary(predictDT_tfidf)

prp(DT_model_tfidf) #tree visualization

printcp(DT_model_tfidf)

confusionMatrix(table(test_tfidf$class, predictDT_tfidf)) #evaluation of confusion matrix
```

Second attempt (10-fold validation)
```{r}
set.seed(481)
folds_1 <- sample(1:10, size = nrow(m_tfidf), replace = T)
table(folds_1)

CV_lda_1 <- lapply(1:10, function(x){ 
  model_1 <- lda(class ~., m_tfidf[folds_1!= x, ])
  preds_1 <- predict(model_1,  m_tfidf[folds_1 == x, ])$class
  return(data.frame(preds_1, real = m_tfidf$class[folds_1 == x]))
})

CV_lda_1 <- do.call(rbind, CV_lda_1)

confusionMatrix(CV_lda_1$preds, CV_lda_1$real)

set.seed(482)
folds_2 <- sample(1:10, size = nrow(m_tfidf), replace = T)
table(folds_2)

CV_lda_2 <- lapply(1:10, function(x){ 
  model_2 <- lda(class ~., m_tfidf[folds_2 != x, ])
  preds_2 <- predict(model_2,  m_tfidf[folds_2 == x, ])$class
  return(data.frame(preds_2, real = m_tfidf$class[folds_2 == x]))
})

CV_lda_2 <- do.call(rbind, CV_lda_2)

confusionMatrix(CV_lda_2$preds, CV_lda_2$real)

set.seed(483)
folds_3 <- sample(1:10, size = nrow(m_tfidf), replace = T)
table(folds_3)

CV_lda_3 <- lapply(1:10, function(x){ 
  model_3 <- lda(class ~., m_tfidf[folds_3!= x, ])
  preds_3 <- predict(model_3,  m_tfidf[folds_3 == x, ])$class
  return(data.frame(preds_3, real = m_tfidf$class[folds_3 == x]))
})

CV_lda_3 <- do.call(rbind, CV_lda_3)

confusionMatrix(CV_lda_3$preds, CV_lda_3$real)

set.seed(484)
folds_4 <- sample(1:10, size = nrow(m_tfidf), replace = T)
table(folds_4)

CV_lda_4 <- lapply(1:10, function(x){ 
  model_4 <- lda(class ~., m_tfidf[folds_4 != x, ])
  preds_4 <- predict(model_4,  m_tfidf[folds_4 == x, ])$class
  return(data.frame(preds_4, real = m_tfidf$class[folds_4 == x]))
})

CV_lda_4 <- do.call(rbind, CV_lda_4)

confusionMatrix(CV_lda_4$preds, CV_lda_4$real)

set.seed(485)
folds_5 <- sample(1:10, size = nrow(m_tfidf), replace = T)
table(folds_5)

CV_lda_5 <- lapply(1:10, function(x){ 
  model_5 <- lda(class ~., m_tfidf[folds_5!= x, ])
  preds_5 <- predict(model_5,  m_tfidf[folds_5 == x, ])$class
  return(data.frame(preds_5, real = m_tfidf$class[folds_5 == x]))
})

CV_lda_5 <- do.call(rbind, CV_lda_5)

confusionMatrix(CV_lda_5$preds, CV_lda_5$real)

set.seed(486)
folds_6 <- sample(1:10, size = nrow(m_tfidf), replace = T)
table(folds_6)

CV_lda_6 <- lapply(1:10, function(x){ 
  model_6 <- lda(class ~., m_tfidf[folds_6 != x, ])
  preds_6 <- predict(model_6,  m_tfidf[folds_6 == x, ])$class
  return(data.frame(preds_6, real = m_tfidf$class[folds_6 == x]))
})

CV_lda_6 <- do.call(rbind, CV_lda_6)

confusionMatrix(CV_lda_6$preds, CV_lda_6$real)

set.seed(487)
folds_7 <- sample(1:10, size = nrow(m_tfidf), replace = T)
table(folds_7)

CV_lda_7 <- lapply(1:10, function(x){ 
  model_7 <- lda(class ~., m_tfidf[folds_7!= x, ])
  preds_7 <- predict(model_7,  m_tfidf[folds_7 == x, ])$class
  return(data.frame(preds_7, real = m_tfidf$class[folds_7 == x]))
})

CV_lda_7 <- do.call(rbind, CV_lda_7)

confusionMatrix(CV_lda_7$preds, CV_lda_7$real)

set.seed(488)
folds_8 <- sample(1:10, size = nrow(m_tfidf), replace = T)
table(folds_8)

CV_lda_8 <- lapply(1:10, function(x){ 
  model_8 <- lda(class ~., m_tfidf[folds_8 != x, ])
  preds_8 <- predict(model_8,  m_tfidf[folds_8 == x, ])$class
  return(data.frame(preds_8, real = m_tfidf$class[folds_8 == x]))
})

CV_lda_8 <- do.call(rbind, CV_lda_8)

confusionMatrix(CV_lda_8$preds, CV_lda_8$real)

set.seed(489)
folds_9 <- sample(1:10, size = nrow(m_tfidf), replace = T)
table(folds_9)

CV_lda_9 <- lapply(1:10, function(x){ 
  model_9 <- lda(class ~., m_tfidf[folds_9!= x, ])
  preds_9 <- predict(model_9,  m_tfidf[folds_9 == x, ])$class
  return(data.frame(preds_9, real = m_tfidf$class[folds_9 == x]))
})

CV_lda_9 <- do.call(rbind, CV_lda_9)

confusionMatrix(CV_lda_9$preds, CV_lda_9$real)

set.seed(410)
folds_10 <- sample(1:10, size = nrow(m_tfidf), replace = T)
table(folds_10)

CV_lda_10 <- lapply(1:10, function(x){ 
  model_10 <- lda(class ~., m_tfidf[folds_10 != x, ])
  preds_10 <- predict(model_10,  m_tfidf[folds_10 == x, ])$class
  return(data.frame(preds_10, real = m_tfidf$class[folds_10 == x]))
})

CV_lda_10 <- do.call(rbind, CV_lda_10)

confusionMatrix(CV_lda_10$preds, CV_lda_10$real)
```

#### Bag-of-Words

```{r}
m_bow <- as.data.frame(tweets_m_bow)
colnames(m_bow) <- make.names(colnames(m_bow))
m_bow$class = twitter_balanced$airline_sentiment
```

First attempt (no repetition)
```{r}
set.seed(989)
split_bow = sample.split(m_bow$class, SplitRatio = 0.7)
train_bow = subset(m_bow, split_bow == TRUE)
test_bow = subset(m_bow, split_bow == FALSE)

DT_model_bow <- rpart(class~., data = train_bow)

predictDT_bow <- predict(DT_model_bow, newdata = test_bow, type = "class")

prp(DT_model_bow) #tree visualization

confusionMatrix(table(test_bow$class, predictDT_bow)) #evaluation of confusion matrix
```

Second attempt (k-fold validation)
```{r}
set.seed(701)
folds_bow_1 <- sample(1:10, size = nrow(m_bow), replace = T)
table(folds_bow_1)

CV_lda_bow_1 <- lapply(1:10, function(x){ 
  model_bow_1 <- lda(class ~., m_bow[folds_bow_1!= x, ])
  preds_bow_1 <- predict(model_bow_1,  m_bow[folds_bow_1 == x, ])$class
  return(data.frame(preds_bow_1, real = m_bow$class[folds_bow_1 == x]))
})

CV_lda_bow_1 <- do.call(rbind, CV_lda_bow_1)

confusionMatrix(CV_lda_bow_1$preds, CV_lda_bow_1$real)

set.seed(702)
folds_bow_2 <- sample(1:10, size = nrow(m_bow), replace = T)
table(folds_bow_2)

CV_lda_bow_2 <- lapply(1:10, function(x){ 
  model_bow_2 <- lda(class ~., m_bow[folds_bow_2 != x, ])
  preds_bow_2 <- predict(model_bow_2,  m_bow[folds_bow_2 == x, ])$class
  return(data.frame(preds_bow_2, real = m_bow$class[folds_bow_2 == x]))
})

CV_lda_bow_2 <- do.call(rbind, CV_lda_bow_2)

confusionMatrix(CV_lda_bow_2$preds, CV_lda_bow_2$real)

set.seed(703)
folds_bow_3 <- sample(1:10, size = nrow(m_bow), replace = T)
table(folds_bow_3)

CV_lda_bow_3 <- lapply(1:10, function(x){ 
  model_bow_3 <- lda(class ~., m_bow[folds_bow_3!= x, ])
  preds_bow_3 <- predict(model_bow_3,  m_bow[folds_bow_3 == x, ])$class
  return(data.frame(preds_bow_3, real = m_bow$class[folds_bow_3 == x]))
})

CV_lda_bow_3 <- do.call(rbind, CV_lda_bow_3)

confusionMatrix(CV_lda_bow_3$preds, CV_lda_bow_3$real)

set.seed(704)
folds_bow_4 <- sample(1:10, size = nrow(m_bow), replace = T)
table(folds_bow_4)

CV_lda_bow_4 <- lapply(1:10, function(x){ 
  model_bow_4 <- lda(class ~., m_bow[folds_bow_4 != x, ])
  preds_bow_4 <- predict(model_bow_4,  m_bow[folds_bow_4 == x, ])$class
  return(data.frame(preds_bow_4, real = m_bow$class[folds_bow_4 == x]))
})

CV_lda_bow_4 <- do.call(rbind, CV_lda_bow_4)

confusionMatrix(CV_lda_bow_4$preds, CV_lda_bow_4$real)

set.seed(705)
folds_bow_5 <- sample(1:10, size = nrow(m_bow), replace = T)
table(folds_bow_5)

CV_lda_bow_5 <- lapply(1:10, function(x){ 
  model_bow_5 <- lda(class ~., m_bow[folds_bow_5!= x, ])
  preds_bow_5 <- predict(model_bow_5,  m_bow[folds_bow_5 == x, ])$class
  return(data.frame(preds_bow_5, real = m_bow$class[folds_bow_5 == x]))
})

CV_lda_bow_5 <- do.call(rbind, CV_lda_bow_5)

confusionMatrix(CV_lda_bow_5$preds, CV_lda_bow_5$real)

set.seed(706)
folds_bow_6 <- sample(1:10, size = nrow(m_bow), replace = T)
table(folds_bow_6)

CV_lda_bow_6 <- lapply(1:10, function(x){ 
  model_bow_6 <- lda(class ~., m_bow[folds_bow_6 != x, ])
  preds_bow_6 <- predict(model_bow_6,  m_bow[folds_bow_6 == x, ])$class
  return(data.frame(preds_bow_6, real = m_bow$class[folds_bow_6 == x]))
})

CV_lda_bow_6 <- do.call(rbind, CV_lda_bow_6)

confusionMatrix(CV_lda_bow_6$preds, CV_lda_bow_6$real)

set.seed(707)
folds_bow_7 <- sample(1:10, size = nrow(m_bow), replace = T)
table(folds_bow_7)

CV_lda_bow_7 <- lapply(1:10, function(x){ 
  model_bow_7 <- lda(class ~., m_bow[folds_bow_7!= x, ])
  preds_bow_7 <- predict(model_bow_7,  m_bow[folds_bow_7 == x, ])$class
  return(data.frame(preds_bow_7, real = m_bow$class[folds_bow_7 == x]))
})

CV_lda_bow_7 <- do.call(rbind, CV_lda_bow_7)

confusionMatrix(CV_lda_bow_7$preds, CV_lda_bow_7$real)

set.seed(708)
folds_bow_8 <- sample(1:10, size = nrow(m_bow), replace = T)
table(folds_bow_8)

CV_lda_bow_8 <- lapply(1:10, function(x){ 
  model_bow_8 <- lda(class ~., m_bow[folds_bow_8 != x, ])
  preds_bow_8 <- predict(model_bow_8,  m_bow[folds_bow_8 == x, ])$class
  return(data.frame(preds_bow_8, real = m_bow$class[folds_bow_8 == x]))
})

CV_lda_bow_8 <- do.call(rbind, CV_lda_bow_8)

confusionMatrix(CV_lda_bow_8$preds, CV_lda_bow_8$real)

set.seed(709)
folds_bow_9 <- sample(1:10, size = nrow(m_bow), replace = T)
table(folds_bow_9)

CV_lda_bow_9 <- lapply(1:10, function(x){ 
  model_bow_9 <- lda(class ~., m_bow[folds_bow_9!= x, ])
  preds_bow_9 <- predict(model_bow_9,  m_bow[folds_bow_9 == x, ])$class
  return(data.frame(preds_bow_9, real = m_bow$class[folds_bow_9 == x]))
})

CV_lda_bow_9 <- do.call(rbind, CV_lda_bow_9)

confusionMatrix(CV_lda_bow_9$preds, CV_lda_bow_9$real)

set.seed(800)
folds_bow_10 <- sample(1:10, size = nrow(m_bow), replace = T)
table(folds_bow_10)

CV_lda_bow_10 <- lapply(1:10, function(x){ 
  model_bow_10 <- lda(class ~., m_bow[folds_bow_10 != x, ])
  preds_bow_10 <- predict(model_bow_10,  m_bow[folds_bow_10 == x, ])$class
  return(data.frame(preds_bow_10, real = m_bow$class[folds_bow_10 == x]))
})

CV_lda_bow_10 <- do.call(rbind, CV_lda_bow_10)

confusionMatrix(CV_lda_bow_10$preds, CV_lda_bow_10$real)
```

#### Unigram

```{r}
m_uni <- as.data.frame(tweets_m_uni)
colnames(m_uni) <- make.names(colnames(m_uni))
m_uni$class = twitter_balanced$airline_sentiment
```

First attempt (no repetition)
```{r}
set.seed(825)
split_uni = sample.split(m_uni$class, SplitRatio = 0.7)
train_uni = subset(m_uni, split_uni == TRUE)
test_uni = subset(m_uni, split_uni == FALSE)

DT_model_uni <- rpart(class~., data = train_uni)

predictDT_uni <- predict(DT_model_uni, newdata = test_uni, type = "class")

prp(DT_model_uni) #tree visualization

confusionMatrix(table(test_uni$class, predictDT_uni)) #evaluation of confusion matrix
```

Second attempt (10-fold validation)
```{r}
set.seed(321)
folds_uni_1 <- sample(1:10, size = nrow(m_uni), replace = T)
table(folds_uni_1)

CV_lda_uni_1 <- lapply(1:10, function(x){ 
  model_uni_1 <- lda(class ~., m_uni[folds_uni_1!= x, ])
  preds_uni_1 <- predict(model_uni_1,  m_uni[folds_uni_1 == x, ])$class
  return(data.frame(preds_uni_1, real = m_uni$class[folds_uni_1 == x]))
})

CV_lda_uni_1 <- do.call(rbind, CV_lda_uni_1)

confusionMatrix(CV_lda_uni_1$preds, CV_lda_uni_1$real)

set.seed(322)
folds_uni_2 <- sample(1:10, size = nrow(m_uni), replace = T)
table(folds_uni_2)

CV_lda_uni_2 <- lapply(1:10, function(x){ 
  model_uni_2 <- lda(class ~., m_uni[folds_uni_2 != x, ])
  preds_uni_2 <- predict(model_uni_2,  m_uni[folds_uni_2 == x, ])$class
  return(data.frame(preds_uni_2, real = m_uni$class[folds_uni_2 == x]))
})

CV_lda_uni_2 <- do.call(rbind, CV_lda_uni_2)

confusionMatrix(CV_lda_uni_2$preds, CV_lda_uni_2$real)

set.seed(323)
folds_uni_3 <- sample(1:10, size = nrow(m_uni), replace = T)
table(folds_uni_3)

CV_lda_uni_3 <- lapply(1:10, function(x){ 
  model_uni_3 <- lda(class ~., m_uni[folds_uni_3!= x, ])
  preds_uni_3 <- predict(model_uni_3,  m_uni[folds_uni_3 == x, ])$class
  return(data.frame(preds_uni_3, real = m_uni$class[folds_uni_3 == x]))
})

CV_lda_uni_3 <- do.call(rbind, CV_lda_uni_3)

confusionMatrix(CV_lda_uni_3$preds, CV_lda_uni_3$real)

set.seed(324)
folds_uni_4 <- sample(1:10, size = nrow(m_uni), replace = T)
table(folds_uni_4)

CV_lda_uni_4 <- lapply(1:10, function(x){ 
  model_uni_4 <- lda(class ~., m_uni[folds_uni_4 != x, ])
  preds_uni_4 <- predict(model_uni_4,  m_uni[folds_uni_4 == x, ])$class
  return(data.frame(preds_uni_4, real = m_uni$class[folds_uni_4 == x]))
})

CV_lda_uni_4 <- do.call(rbind, CV_lda_uni_4)

confusionMatrix(CV_lda_uni_4$preds, CV_lda_uni_4$real)

set.seed(325)
folds_uni_5 <- sample(1:10, size = nrow(m_uni), replace = T)
table(folds_uni_5)

CV_lda_uni_5 <- lapply(1:10, function(x){ 
  model_uni_5 <- lda(class ~., m_uni[folds_uni_5!= x, ])
  preds_uni_5 <- predict(model_uni_5,  m_uni[folds_uni_5 == x, ])$class
  return(data.frame(preds_uni_5, real = m_uni$class[folds_uni_5 == x]))
})

CV_lda_uni_5 <- do.call(rbind, CV_lda_uni_5)

confusionMatrix(CV_lda_uni_5$preds, CV_lda_uni_5$real)

set.seed(326)
folds_uni_6 <- sample(1:10, size = nrow(m_uni), replace = T)
table(folds_uni_6)

CV_lda_uni_6 <- lapply(1:10, function(x){ 
  model_uni_6 <- lda(class ~., m_uni[folds_uni_6 != x, ])
  preds_uni_6 <- predict(model_uni_6,  m_uni[folds_uni_6 == x, ])$class
  return(data.frame(preds_uni_6, real = m_uni$class[folds_uni_6 == x]))
})

CV_lda_uni_6 <- do.call(rbind, CV_lda_uni_6)

confusionMatrix(CV_lda_uni_6$preds, CV_lda_uni_6$real)

set.seed(327)
folds_uni_7 <- sample(1:10, size = nrow(m_uni), replace = T)
table(folds_uni_7)

CV_lda_uni_7 <- lapply(1:10, function(x){ 
  model_uni_7 <- lda(class ~., m_uni[folds_uni_7!= x, ])
  preds_uni_7 <- predict(model_uni_7,  m_uni[folds_uni_7 == x, ])$class
  return(data.frame(preds_uni_7, real = m_uni$class[folds_uni_7 == x]))
})

CV_lda_uni_7 <- do.call(rbind, CV_lda_uni_7)

confusionMatrix(CV_lda_uni_7$preds, CV_lda_uni_7$real)

set.seed(328)
folds_uni_8 <- sample(1:10, size = nrow(m_uni), replace = T)
table(folds_uni_8)

CV_lda_uni_8 <- lapply(1:10, function(x){ 
  model_uni_8 <- lda(class ~., m_uni[folds_uni_8 != x, ])
  preds_uni_8 <- predict(model_uni_8,  m_uni[folds_uni_8 == x, ])$class
  return(data.frame(preds_uni_8, real = m_uni$class[folds_uni_8 == x]))
})

CV_lda_uni_8 <- do.call(rbind, CV_lda_uni_8)

confusionMatrix(CV_lda_uni_8$preds, CV_lda_uni_8$real)

set.seed(329)
folds_uni_9 <- sample(1:10, size = nrow(m_uni), replace = T)
table(folds_uni_9)

CV_lda_uni_9 <- lapply(1:10, function(x){ 
  model_uni_9 <- lda(class ~., m_uni[folds_uni_9!= x, ])
  preds_uni_9 <- predict(model_uni_9,  m_uni[folds_uni_9 == x, ])$class
  return(data.frame(preds_uni_9, real = m_uni$class[folds_uni_9 == x]))
})

CV_lda_uni_9 <- do.call(rbind, CV_lda_uni_9)

confusionMatrix(CV_lda_uni_9$preds, CV_lda_uni_9$real)

set.seed(400)
folds_uni_10 <- sample(1:10, size = nrow(m_uni), replace = T)
table(folds_uni_10)

CV_lda_uni_10 <- lapply(1:10, function(x){ 
  model_uni_10 <- lda(class ~., m_uni[folds_uni_10 != x, ])
  preds_uni_10 <- predict(model_uni_10,  m_uni[folds_uni_10 == x, ])$class
  return(data.frame(preds_uni_10, real = m_uni$class[folds_uni_10 == x]))
})

CV_lda_uni_10 <- do.call(rbind, CV_lda_uni_10)

confusionMatrix(CV_lda_uni_10$preds, CV_lda_uni_10$real)
```

From the decision tree, the TF-IDF model resulted in the highest accuracy at 59.35%; however, this is very close to the other models (Bag-of-Words, Unigrams). For all three algorithm-model combination, the word "thank" is the root node followed by "hour" and "great". 

### Random Forest

Apply the Random Forest algorithm to the three models: TF-IDF, Bag-of-Words, and Unigram. 

#### TF-IDF

```{r}
RF_model_tfidf <- randomForest(class~., data = train_tfidf)

predictRF_tfidf <- predict(RF_model_tfidf, newdata = test_tfidf)

confusionMatrix(table(test_tfidf$class, predictRF_tfidf)) #evaluation of confusion matrix 
```

#### Bag-of-Words

```{r}
RF_model_bow <- randomForest(class~., data = train_bow)

predictRF_bow <- predict(RF_model_bow, newdata = test_bow)

confusionMatrix(table(test_bow$class, predictRF_bow)) #evaluation of confusion matrix 
```

#### Unigram

```{r}
RF_model_uni <- randomForest(class~., data = train_uni)

predictRF_uni <- predict(RF_model_uni, newdata = test_uni)

confusionMatrix(table(test_uni$class, predictRF_uni)) #evaluation of confusion matrix 
```

### Naive Bayes

Apply the Naive Bayes algorithm to the three models: TF-IDF, Bag-of-Words, and Unigram. 

#### TF-IDF

```{r}
NB_model_tfidf <- naiveBayes(class~., data = train_tfidf)

predictNB_tfidf <- predict(NB_model_tfidf, newdata = test_tfidf)

confusionMatrix(table(test_tfidf$class, predictNB_tfidf)) #evaluation of confusion matrix
```

#### Bag-of-Words

```{r}
NB_model_bow <- naiveBayes(class~., data = train_bow)

predictNB_bow <- predict(NB_model_bow, newdata = test_bow)

confusionMatrix(table(test_bow$class, predictNB_bow)) #evaluation of confusion matrix
```

#### Unigram

```{r}
NB_model_uni <- naiveBayes(class~., data = train_uni)

predictNB_uni <- predict(NB_model_uni, newdata = test_uni)

confusionMatrix(table(test_uni$class, predictNB_uni)) #evaluation of confusion matrix
```